Embedding Dimension of Polyhedral Losses

### 2020-01-31 colt-20/embed-dim.tex
small wording changes

\paragraph{Properties of distributions.}
In order to study calibration of surrogates (Definition~\ref{def:calibration} below), it is useful to formalize the following functions that describe the reports that minimize expected $\ell$-loss and $L$-loss under $p$.
%% Bo: since we use \gamma and \R for \ell, and we use \Gamma and \reals^d for L,
%% I figured generic properties could use notation g and R for f
\begin{definition}[(Finite) property, level set, elicited]
  A \emph{property} is a function $g: \simplex \to 2^{R}$ for some set $R$ where $g(p) \neq \emptyset$ for all $p$.
  It is \emph{finite} if $|R| < \infty$.
  The \emph{level set} of $r \in R$ is the set $g_r = \{p \in \simplex : r \in g(p)\}$.
  A loss function $f: R \to \nonnegreals^{\Y}$ \emph{elicits} $g$ if $g(p) = \argmin_{r \in R} \inprod{p}{f(r)}$.
  We assume all properties are non-redundant, i.e. each $g_r$ contains a distribution $p$ not in any other $g_{r'}$.\jessiet{$R$ or $\R$?}
\end{definition}
For example, $0-1$ loss elicits the mode, which is formalized as a property on report space $R = \Y$ via $g(p) = \argmax_{y \in \Y} p(y)$.
Notice in this example $g(p)$ is a singleton set when the mode of $p$ is unique and e.g. is $\Y$ when $p$ is uniform.
%Another example of a finite property is ranking, in which $\R$ is the set of permutations of $\Y$, and $g(p)$ consists of all permutations that order $\Y$ from most to least probable (i.e. ties must be broken in all possible ways).
In this paper we write $\gamma$ for the property elicited by $\ell$ with report space $\R$; and we use $\Gamma$ for the property elicited by $L$ with report space $\reals^d$.

\paragraph{Embeddings.}
%We specifically study polyhedral functions in part because they have high-dimensional subgradient sets at points of nondifferentiability, which will be of interest to us later.
Polyhedral losses are motivated partly because they correspond to a natural surrogate construction technique: embedding the discrete predictions $\R$ as points in $\reals^d$ and linearly interpolating between embedded reports.
The key condition required on the embedding and surrogate loss $L$ is that a discrete prediction $r \in \R$ should minimize expected $\ell$-loss if and only if its embedding point minimizes expected $L$-loss.
%
\begin{definition}[Embedding of a loss]
  A loss $L: \reals^d \to \nonnegreals^{\Y}$ \emph{embeds} a loss $\ell: \R \to \nonnegreals^{\Y}$ \emph{in dimension $d$} with \emph{embedding function $\varphi: \R \to \reals^d$} if: (i) $\varphi$ is injective; (ii) for all $r \in \R$, $\ell(r) = L(\varphi(r))$; and (iii) for all $r \in \R$, $\gamma_r = \Gamma_{\varphi(r)}$.
  We simply say $L$ \emph{embeds} $\ell$ if some such $\varphi$ exists.
\end{definition}
%
Note that embedding does not immediately imply a key desirable property, consistency, which is equivalent to calibration in finite-outcome settings.
Informally, calibration means that if one minimizes the expected surrogate loss arbitrarily well over $\reals^d$, then applies the link function to obtain a discrete prediction $r$, then $r$ exactly minimizes $\ell$-loss.
\begin{definition}[Calibrated surrogate]\label{def:calibration}
  A surrogate loss $L: \reals^d \to \nonnegreals^{\Y}$ and \emph{link function} $\psi: \reals^d \to \R$ are \emph{calibrated} for a loss $\ell: \R \to \nonnegreals^{\Y}$ if for all $p \in \simplex$,
    \[ \inf_{u \in \reals^d ~:~ \psi(u) \not\in \gamma(p)} \inprod{p}{L(u)}  > \inf_{u \in \reals^d} \inprod{p}{L(u)}  .\]
  We simply say $L$ can be \emph{calibrated} to $\ell$ if such a link function exists.
\end{definition}
As we alluded to above, calibration is not actually the trait we desire of a surrogate; rather, we seek consistency.
\cite{tewari2007consistency} show that, in a finite-outcome setting, the presence of calibration and consistency are equivalent conditions.
By consistency, we mean the standard notion where the expected loss over $\X \times \Y$ for a sequence of hypotheses $h_m$ converging to the optimal expected surrogate loss implies the expected discrete loss over the linked hypotheses $\psi \circ h_m$ converges to the optimal discrete loss.
The equivalence of these conditions allows us to focus on calibration, so we can abstract away the input space $\X$ and simply focus on probability distributions over the outcomes $\Y$.

@@

  Moreover, when these conditions hold, the loss $L:\reals\to\reals^\Y$ embeds $\ell$ with $\varphi:\R\to\reals$, \jessiet{Have been using $\phi$ instead of $\varphi$ for embeddings in this paper so far. I have a slight preference for $\varphi$ since it matches the last paper, but I don't care too much... just need to pick one.}\raft{Why switch to $\phi$ if you prefer $\varphi$?}
  where
  \begin{align*}
    \varphi(r_i) &= \sum_{j=1}^{i-1} 1/\Lambda_j~, \quad \text{(where $\varphi(r_1)=0$)}
    \\
    L(u) &= \begin{cases}
      \ell(r_1) - u K \ones & u \leq \varphi(r_1) = 0 \\
      \ell(r_i) + \Lambda_i \cdot (u - \varphi(r_i)) \cdot  (\ell(r_{i+1})-\ell(r_i)) & u \in [\varphi(r_i),\varphi(r_{i+1})] \\
      \ell(r_k) + \Lambda_{k-1} \cdot (u - \varphi(r_k))\cdot K \ones & u \geq \varphi(r_k)
    \end{cases}~,
  \end{align*}
  \jessiet{In the cases: what if $\varphi(r_1) < 0$?}\raft{I set $\varphi(r_1)=0$; updated to something hopefully clearer}
  where $\lambda(i) = \min(\lambda^+(i),\max(\lambda^-(i),1))$, $\Lambda_i := \prod_{j=2}^i \lambda(j)$, $\Lambda_1=1$, and $K = \max_{i\in\{2,\ldots,k\},y\in\Y} |v(i)_y|$.
\end{theorem}
Note that a calibrated surrogate should also specify a link function $\psi: \reals^d \to \R$.
We have been omitting it because of the guarantee of Theorem \ref{thm:embed-iff-poly}, from \cite{finocchiaro2019embedding}, which guarantees its existence.
But in the case of $d=1$, we observe that one can simply use the midpoints between the embedding points as cutoffs for the link function, i.e. $\psi(u) = \argmin_{r \in \R} |u - \phi(r)|$, breaking ties arbitrarily.

@@

\begin{theorem} \label{thm:general-char}
  Let $\ell: \R \to \reals_{\geq 0}^{\Y}$ be a discrete loss with, for each $r \in \R$, $\gamma_r = \{p : r \in \argmin_{r'} \inprod{p}{\ell(r)}\}$.
  Then $\ell$ is $d$-embeddable if and only if there exist polytopes $\T = \{T_y \subseteq \reals^d : y \in \Y\}$ for each $r \in \R$ such that both of the following hold:
  \begin{enumerate}
    \item (Optimality) We have $D(\T) = \gamma_r$.
    \item (Monotonicity) There exists an injective embedding function $\varphi : \R \to \reals^d$ and loss functions $\{L_y : \reals^d \to \reals^\Y_{\geq 0}\}_{y \in \Y}$ such that for all $y \in \Y$, we have $T_y = \partial L_y(\varphi(r))$ and for all $r \in \R$, we have $L_y(\varphi(r)) = \ell_y(r)$.
  \end{enumerate}
\end{theorem}
\begin{proof}
  \jessie{An attempt:}
  $(\implies)$ If $\ell$ is $d$-embeddable, say that $L:\reals^d \to \reals^\Y_{\geq 0}$ embeds $\ell$.
  Take $T_y$ to be $\partial L(\varphi(r))_y$ for a fixed $r \in \R$ and $y \in \Y$.
  Clearly this set of polytopes $\T$ satisfies the monotonicity condition.
  To see the optimality condition is satisfied, consider that $\gamma_r = \Gamma_{\varphi(r)} = \{p \in \simplex : \vec 0 \in \oplus_p \partial L(\varphi(r))_y \} = \{p : \vec 0 \in \oplus_p T_y \} = D(\T)$.

@@

Thus, although feasible subspace dimension applies to any convex surrogate, the relation to our techniques is an interesting future direction.
The advantage of our approach for generic $d$ is that our optimality conditions and quadratic feasibility program consider the structure of the entire level set $\gamma_r$, rather than just a single witness point.
This allows us to prove lower bounds on the abstain loss for $\alpha \leq 1/2$, which feasible subspace dimension cannot.
Proving the following conjecture would even more closely relate the two techniques (a disproof would also be extremely interesting):
\begin{conjecture}\label{conj:embed-equals-ccdim}
  If a discrete loss $\ell$ has a $d$-dimensional calibrated convex surrogate, then it is $d$-embeddable (i.e. has a $d$-dimensional calibrated polyhedral surrogate).
  In other words, embedding dimension always equals convex calibration dimension~\citep{ramaswamy2016convex}.
\end{conjecture}
Our results show the conjecture is true when $d=1$.

\paragraph{Future Work.}
There are a few threads of future work: the first is to utilize monotonicity to see if we can construct even tighter lower bounds on embedding dimension.
Second, we hope to understand when, if ever, embedding dimension is not equal to convex calibration dimension, as stated in Conjecture~\ref{conj:embed-equals-ccdim}.
Moreover, the restriction that we are calibrated over the entire simplex may be tighter than necessary in some contexts, and would be useful to understand the tradeoff between calibration and dimension of a surrogate loss.
For one example where we can reduce surrogate dimension with a low-entropy assumption on the simplex, see~\cite[Example 6]{agarwal2015consistent}.

>>>>

\paragraph{Properties of distributions.}
In order to study calibration of surrogates (Definition~\ref{def:calibration} below), it is useful to formalize the following functions that describe the reports that minimize expected $\ell$-loss and $L$-loss under $p$.
%% Bo: since we use \gamma and \R for \ell, and we use \Gamma and \reals^d for L,
%% I figured generic properties could use notation g and R for f
\begin{definition}[(Finite) property, level set, elicited]
  A \emph{property} is a function $g: \simplex \to 2^{R}$ for some set $R$ where $g(p) \neq \emptyset$ for all $p$.
  It is \emph{finite} if $|R| < \infty$.
  The \emph{level set} of $r \in R$ is the set $g_r = \{p \in \simplex : r \in g(p)\}$.
  A loss function $f: R \to \nonnegreals^{\Y}$ \emph{elicits} $g$ if $g(p) = \argmin_{r \in R} \inprod{p}{f(r)}$.
  We assume all properties are non-redundant, i.e. each $g_r$ contains a distribution $p$ not in any other $g_{r'}$.
\end{definition}
For example, $0-1$ loss elicits the mode, which is formalized as a property on report space $R = \Y$ via $g(p) = \argmax_{y \in \Y} p(y)$.
Notice in this example $g(p)$ is a singleton set when the mode of $p$ is unique and e.g. is $\Y$ when $p$ is uniform.
%Another example of a finite property is ranking, in which $\R$ is the set of permutations of $\Y$, and $g(p)$ consists of all permutations that order $\Y$ from most to least probable (i.e. ties must be broken in all possible ways).
In this paper we write $\gamma$ for the property elicited by $\ell$ with report space $\R$; and we use $\Gamma$ for the property elicited by $L$ with report space $\reals^d$.

\paragraph{Embeddings.}
%We specifically study polyhedral functions in part because they have high-dimensional subgradient sets at points of nondifferentiability, which will be of interest to us later.
Polyhedral losses are motivated partly because they correspond to a natural surrogate construction technique: embedding the discrete predictions $\R$ as points in $\reals^d$ and linearly interpolating between embedded reports.
The key condition required on the embedding and surrogate loss $L$ is that a discrete prediction $r \in \R$ should minimize expected $\ell$-loss if and only if its embedding point minimizes expected $L$-loss.
%
\begin{definition}[Embedding of a loss]
  A loss $L: \reals^d \to \nonnegreals^{\Y}$ \emph{embeds} a loss $\ell: \R \to \nonnegreals^{\Y}$ \emph{in dimension $d$} with \emph{embedding function $\varphi: \R \to \reals^d$} if: (i) $\varphi$ is injective; (ii) for all $r \in \R$, $\ell(r) = L(\varphi(r))$; and (iii) for all $r \in \R$, $\gamma_r = \Gamma_{\varphi(r)}$.
  We simply say $L$ \emph{embeds} $\ell$ if some such $\phi$ exists.
\end{definition}
%
Note that embedding does not immediately imply a key desirable property, consistency, which is equivalent to calibration in finite-outcome settings.
Informally, calibration means that if one minimizes the expected surrogate loss arbitrarily well over $\reals^d$, then applies the link function to obtain a discrete prediction $r$, then $r$ exactly minimizes $\ell$-loss.
\begin{definition}[Calibrated surrogate]\label{def:calibration}
  A surrogate loss $L: \reals^d \to \nonnegreals^{\Y}$ and \emph{link function} $\psi: \reals^d \to \R$ are \emph{calibrated} for a loss $\ell: \R \to \nonnegreals^{\Y}$ if for all $p \in \simplex$,
    \[ \inf_{u \in \reals^d ~:~ \psi(u) \not\in \gamma(p)} \inprod{p}{L(u)}  > \inf_{u \in \reals^d} \inprod{p}{L(u)}  .\]
  We simply say $L$ can be \emph{calibrated} to $\ell$ if such a link function exists.
\end{definition}
As we alluded to above, calibration is not actually the trait we desire of a surrogate; rather, we seek consistency.
\cite{tewari2007consistency} show that, in a finite-outcome setting, the presence of calibration and consistency are equivalent conditions.
By consistency, we mean the standard notion where the expected loss over $\X \times \Y$ for a sequence of hypotheses $h_m$ converging to the optimal expected surrogate loss implies the expected discrete loss over the linked hypotheses $\psi \circ h_m$ converges to the optimal discrete loss.
The equivalence of these conditions allows us to focus on calibration, so we can abstract away the input space $\X$ and simply focus on probability distributions over the outcomes $\Y$.

@@

  Moreover, when these conditions hold, the loss $L:\reals\to\reals^\Y$ embeds $\ell$ with $\varphi:\R\to\reals$, \jessiet{Have been using $\phi$ instead of $\varphi$ for embeddings in this paper so far. I have a slight preference for $\varphi$ since it matches the last paper, but I don't care too much... just need to pick one.}\raft{Why switch to $\phi$ if you prefer $\varphi$?}
  where
  \begin{align*}
    \varphi(r_i) &= \sum_{j=2}^i 1/\Lambda_j~, \quad \text{(where $\varphi(r_1)=0$)}
    \\
    L(u) &= \begin{cases}
      \ell(r_1) - u K \ones & u \leq \varphi(r_1) = 0 \\
      \ell(r_i) + \Lambda_i \cdot (u - \varphi(r_i)) \cdot  (\ell(r_{i+1})-\ell(r_i)) & u \in [\varphi(r_i),\varphi(r_{i+1})] \\
      \ell(r_k) + \Lambda_{k-1} \cdot (u - \varphi(r_k))\cdot K \ones & u \geq \varphi(r_k)
    \end{cases}~,
  \end{align*}
  \jessiet{In the cases: what if $\varphi(r_1) < 0$?}\raft{I set $\varphi(r_1)=0$; updated to something hopefully clearer}
  where $\lambda(i) = \min(\lambda^+(i),\max(\lambda^-(i),1))$, $\Lambda_i := \prod_{j=2}^i \lambda(j)$, $\Lambda_1=1$, and $K = \max_{i\in\{2,\ldots,k\},y\in\Y} |v(i)_y|$.
\end{theorem}
Note that a calibrated surrogate should also specify a link function $\psi: \reals^d \to \R$.
We have been omitting it because of the guarantee of Theorem \ref{thm:embed-iff-poly}, from \cite{finocchiaro2019embedding}, which guarantees its existence.
But in the case of $d=1$, we observe that one can simply use the midpoints between the embedding points as cutoffs for the link function, i.e. $\psi(u) = \argmin_{r \in \R} |u - \phi(r)|$, breaking ties arbitrarily.

@@

\begin{theorem} \label{thm:general-char}
  Let $\ell: \R \to \reals_{\geq 0}^{\Y}$ be a discrete loss with, for each $r \in \R$, $\gamma_r = \{p : r \in \argmin_{r'} \inprod{p}{\ell(r)}\}$.
  Then $\ell$ is $d$-embeddable if and only if there exists a collection of polytopes $\T^r = \{T^r_y : y \in \Y\}$ for each $r \in \R$ such that both of the following hold:
  \begin{enumerate}
    \item (Optimality) For each $r$, we have $D(\T^r) = \gamma_r$.
    \item (Monotonicity) There exists an injective embedding function $\varphi : \R \to \reals^d$ and loss functions $\{L_y : \reals^d \to \reals^\Y_{\geq 0}\}_{y \in \Y}$ such that for all $r \in \R$ and $y \in \Y$, we have $T^r_y = \partial L_y(\varphi(r))$ and for all $r \in \R$, we have $L_y(\varphi(r)) = \ell_y(r)$.
  \end{enumerate}
\end{theorem}
\begin{proof}
  \jessie{An attempt:}
  $(\implies)$ If $\ell$ is $d$-embeddable, say that $L:\reals^d \to \reals^\Y_{\geq 0}$ embeds $\ell$.
  Take $T_y$ to be $\partial L(\varphi(r))_y$ for a fixed $r \in \R$ and $y \in \Y$.
  Clearly this set of polytopes $\T$ satisfies the monotonicity condition.
  To see the optimality condition is satisfied, consider that $\gamma_r = \Gamma_{\varphi(r)} = \{p \in \simplex : \vec 0 \in \oplus_p \partial L(\varphi(r))_y \} = \{p : \vec 0 \in \oplus_p T_y \} = D(\T)$.

@@

Thus, although feasible subspace dimension applies to any convex surrogate, the relation to our techniques is an interesting future direction.
The advantage of our approach for generic-$d$ is that our optimality conditions and quadratic feasibility program consider the structure of the entire level set $\gamma_r$, rather than just a single point.
This allows us to prove lower bounds on the abstain loss for $\alpha \leq 1/2$, which feasible subspace dimension cannot.
Proving the following conjecture would even more closely relate the two techniques (a disproof would also be extremely interesting):
\begin{conjecture}
  If a discrete loss $\ell$ has a $d$-dimensional calibrated convex surrogate, then it is $d$-embeddable (i.e. has a $d$-dimensional calibrated polyhedral surrogate).
  In other words, embedding dimension always equals convex calibration dimension~\citep{ramaswamy2016convex}.
\end{conjecture}
Our results show the conjecture is true when $d=1$.

\paragraph{Future Work.}
There are a few threads of future work: the first is to utilize monotonicity to see if we can construct even tighter lower bounds on embedding dimension.
Second, we hope to understand when, if ever, embedding dimension is not equal to convex calibration dimension.
Moreover, the restriction that we are calibrated over the entire simplex may be tighter than necessary in some contexts, and would be useful to understand the tradeoff between calibration and dimension of a surrogate loss.
For one example where we can reduce surrogate dimension with a low-entropy assumption on the simplex, see~\cite[Example 6]{agarwal2015consistent}.
